Summer 2024 Report

### 2024-09-06 Relations/Report.tex
ars

\section{09.05}
\begin{enumerate}
  \item done: WN, UN, RP imply SN (classical)
  \item need: omega bounded and RP implies SN (classical)
  \item need: omega bounded and RP implies SN (constructive)
  \item do we need $s_0 \to^* s_c$ in the proof?
  \item is $\omega$-bounded implied by WN and the following confluence property:
  $a \to^* n \in NF$, and $a \to b$ then $b \to^* n $.
  \item Remark before 1.2? (Exercise 1.3.21)
  \item Make the first argument to "being a normal form" implicit?
\end{enumerate}
\end{document}

>>>>

\section{09.05}
\begin{enumerate}
  \item done: WN, UN, RP imply SN (classical)
  \item need: omega bounded and RP implies SN (classical)
  \item need: omega bounded and RP implies SN (constructive)
  \item do we need $s_0 \to^* s_c$ in the proof?
  \item is $\omega$-bounded implied by WN and the following confluence property:
  $a \to^* n \in NF$, and $a \to b$ then $b \to^* n $.
  \item Remark before 1.2? (Exercise 1.3.21)
  \item Make the first argument to "being a normal form" implicit?
\end{enumerate}

\newcommand{\RP}{\mathsf{RP}}
\newcommand{\NF}{\mathsf{NF}}
\newcommand{\UN}{\mathsf{UN}}
\newcommand{\UNto}{\mathsf{UN}{\to}}
\newcommand{\SN}{\mathsf{SN}}
\newcommand{\WN}{\mathsf{WN}}
\newcommand{\from}{\leftarrow}

\subsection*{Additional remarks about Theorem 1.2.3.}
\begin{itemize}
  \item It seems that part (i) of the theorem can be strengthened to only
  assume $\UNto(R)$ instead of $\UN(R)$:
  \[ \WN(R) \land \UNto(R) \to \omega{-}\text{bounded}(R) \tag{i} \]
  \item The analogue of part (ii) of the theorem is unprovable, namely
  \[ \omega{-}\text{bounded}(R) \land \RP(R) \to \SN(R) \tag{ii}\]
  For a countermodel, take the one-element loop $a \to a$.

  (It is both $\omega${-}bounded and satisfies RP.)

  Our hypothesis RP is indeed weaker than ``Inc'' from Terese.
  \item The goal \texttt{iii-lemma :  $WN(R) \to WCR(R) \to \omega{-}\text{bounded}(R)$}
  seems either very difficult or impossible.

  Yet, constructing the bound of a given $R$-increasing sequence is key to the proof of
  (our interpretation of) Theorem 1.2.3(iii), namely that
  \[ WN(R) \land WCR(R) \land RP(R) \to SN(R) \tag{iii} \]
  \item Here is a classical proof of this implication.

  Let $R \subseteq A \times A$, and suppose $R$ satisfies WN, WCR, and RP.

  Suppose $R$ is not SN.  Let $a \in A$ be not SN.  By WN, let $n \in \NF$ be
  such that $a \to^* n$.  By induction on the length of this reduction, the finite sequence
  of terms comprising it must contain a final term with the property that
  it is not SN.  (For $n$ is SN, being a normal form, while $a$ is not.)

  $(\star)$ Let this last non-SN term be denoted $b_0$.

  If each single-step $R$-reduct of $b_0$ is SN (so that $b_0 \to x$ implies $x\in \SN$),
  then $b_0$ would itself be SN --- which it is not, by how it was chosen.

  Thus there exists a $c_0$ with $b_0 \to c_0$, and $c_0 \notin \SN$.

  So we have $c_0 \from b_0 \to m_0$ for some $m_0 \in \SN$. ($m_0$ is the next term after $b_0$ in the reduction $a \to^* n$.)

  By WCR, there exists another element $d_0 \in A$ with $c_0 \to^* d_0 \from^* m_0$.

  Since $m_0 \in \SN$ and this set of terms is closed under reduction, $d_0 \in \SN$ as well.

  Thus $b_0 \to c_0 \to^* d_0$, with $d_0 \in \SN$, and $c_0 \notin \SN$.

  Again, the reduction $c_0 \to^* d_0$ contains a last term that is not in SN.

  Let this term be denoted by $b_1$, and let the next term in the reduction sequence
  to $d_0$ be denoted $m_1$.

  Now repeat the above process starting at $(\star)$, incrementing all subscripts by 1.

  This generates an infinite reduction sequence

  \[b_0 \to c_0 \to^* b_1 \to c_1 \to^* b_2 \to \cdots \]

  Note that infinitely many terms in this sequence are one reduction step
  away from a term in $\SN$.

  Moreover, notice that the normal forms of $m_0$, $d_0$, and $m_1$ are the same
  and equal to $n$, whence the same is true for $m_1$, $d_1$, and $m_2$, etc.

  It follows that, for infinitely many terms in the sequence
  (namely, all the terms $b_i$), we have $b_i \to^* n$.

  Thus, all terms in the sequence reduce to $n$ and so $n$ is an $\omega$-bound
  of the sequence.

  By RP, some term $r$ in the sequence is recurrent.

  Since $r \to^* n$ and $r$ is recurrent, also $n \to^* r$.

  But $n$ is a normal form.  So this sequence is empty and $r=n$.

  But $r$ is a term in an infinite reduction sequence, and reduces to its successor.

  This yields a contradiction.
\end{itemize}
\end{document}
